prod_prep_mergedata

### 2023-12-11 production/prod_analysis.Rmd
update alternativas y readme
* update analisis alternativos
* update link readme

```{r anxiety}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
#load(file = here::here("input/data-proc/df_study1_long_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study1_long_comunas %>%
  select(idencuesta,ola,comuna,just_educ:merit_talent, ahead_family:esc_cit17) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )
# Hipotesis
h1 <- "ola"
h2 <- "ola+merit_effort+merit_talent"
h3 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork"
h4 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork+esc_cit17"
h5 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork+esc_cit17+educ+quintil1+ess+pos_id+edad+sexo"

# A. Anxiety (in_ansied)
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)+(1|comuna)")),data = df_study1)
educ1 <- lmer(formula(paste0("just_educ~",h1,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ2 <- lmer(formula(paste0("just_educ~",h2,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ3 <- lmer(formula(paste0("just_educ~",h3,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ4 <- lmer(formula(paste0("just_educ~",h4,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ5 <- lmer(formula(paste0("just_educ~",h5,"+(1|idencuesta)+(1|comuna)")),data = df_study1)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(educ.null, educ1, educ2, educ3, educ4, educ5),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "ola 2017 <br> <i>(Ref. ola 2016)</i>",
                                      "ola 2018",
                                      "ola 2019",
                                      "ola 2022",
                                      "Merito: esfuerzo",
                                      "Merito: Inteligencia",
                                      "Goahead: Familia",
                                      "Goahead: educacion",
                                      "Goahead: ambicion",
                                      "Goahead: trabajo duro",
                                      "Promedio escolaridad comuna",
                                      "Ed. media <br> <i>(Ref. Ed. básica)</i>",
                                      "Ed. técnica",
                                      "Ed. universitaria",
                                      "Quintil 2 <br> <i>(Ref. Quintil 1)</i>",
                                      "Quintil 3",
                                      "Quintil 4",
                                      "Quintil 5",
                                      "Quintil NA",
                                      "Estatus social subjetivo",
                                      "Pos. política: Centro <br> <i>(Ref. izquierda)</i>",
                                      "Pos. política: Derecha",
                                      "Pos. política: No se identifica",
                                      "Edad 30-49 <br> <i>(Ref. 18-29)</i>",
                                      "Edad 50-64",
                                      "Edad 65 o más",
                                      "Mujer <br> <i>(Ref. Hombre)</i>"
                              )
          #omit.coef = omit
          )
```

>>>>

```{r distribution-educ}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study1_long_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study1_long_comunas %>%
  select(idencuesta,ola,comuna,just_educ:merit_talent, ahead_family:sexo, ginilag22_17, gini_ypch_2022, gini_ypch, prop_univ, gini_gc, prop_pagado, total_mat_pagado, var_simce_comuna, prom_simce_comuna) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )
# Hipotesis
h1 <- "ola"
h2 <- "ola+merit_effort+merit_talent"
h3 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork"
h4 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork+gini_ypch+gini_gc"
h5 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork+gini_ypch+gini_gc+prop_pagado+total_mat_pagado+var_simce_comuna+prom_simce_comuna"
h6 <- "ola+merit_effort+merit_talent+ahead_family+ahead_educ+ahead_ambition+ahead_hardwork+gini_ypch+gini_gc+prop_pagado+total_mat_pagado+var_simce_comuna+prom_simce_comuna+educ+quintil1+ess+pos_id+edad+sexo"

# A. Education distributive justice
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)+(1|comuna)")),data = df_study1)
educ1 <- lmer(formula(paste0("just_educ~",h1,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ2 <- lmer(formula(paste0("just_educ~",h2,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ3 <- lmer(formula(paste0("just_educ~",h3,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ4 <- lmer(formula(paste0("just_educ~",h4,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ5 <- lmer(formula(paste0("just_educ~",h5,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ6 <- lmer(formula(paste0("just_educ~",h6,"+(1|idencuesta)+(1|comuna)")),data = df_study1)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(educ.null, educ1, educ2, educ3, educ4, educ5, educ6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "ola 2017 <br> <i>(Ref. ola 2016)</i>",
                                      "ola 2018",
                                      "ola 2019",
                                      "ola 2022",
                                      "Merito: esfuerzo",
                                      "Merito: Inteligencia",
                                      "Goahead: Familia",
                                      "Goahead: educacion",
                                      "Goahead: ambicion",
                                      "Goahead: trabajo duro",
                                      "Gini comunal",
                                      "Cambio gini 2022-2017",
                                      "Proporción matrícula privada",
                                      "Total matrícula privada",
                                      "Varianza Simce comuna",
                                      "Promedio Simce comuna",
                                      "Ed. media <br> <i>(Ref. Ed. básica)</i>",
                                      "Ed. técnica",
                                      "Ed. universitaria",
                                      "Quintil 2 <br> <i>(Ref. Quintil 1)</i>",
                                      "Quintil 3",
                                      "Quintil 4",
                                      "Quintil 5",
                                      "Quintil NA",
                                      "Estatus social subjetivo",
                                      "Pos. política: Centro <br> <i>(Ref. izquierda)</i>",
                                      "Pos. política: Derecha",
                                      "Pos. política: No se identifica",
                                      "Edad 30-49 <br> <i>(Ref. 18-29)</i>",
                                      "Edad 50-64",
                                      "Edad 65 o más",
                                      "Mujer <br> <i>(Ref. Hombre)</i>"
                              )
          #omit.coef = omit
          )
```

## Contextual analysis

```{r distribution-educ-context}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study1_long_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study1_long_comunas %>%
  select(idencuesta,ola,comuna,just_educ:merit_talent, ahead_family:sexo, ginilag22_17, gini_ypch_2022, gini_ypch, prop_univ, gini_gc, prop_pagado, total_mat_pagado, var_simce_comuna, prom_simce_comuna) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )
# Hipotesis
h1b <- "gini_ypch"
h2b <- "gini_gc"
h3b <- "prop_pagado"
h4b <- "total_mat_pagado"
h5b <- "var_simce_comuna"
h6b <- "prom_simce_comuna"
h7b <- "gini_ypch+gini_gc+prop_pagado+total_mat_pagado+var_simce_comuna+prom_simce_comuna"

# A. Education distributive justice
educ.null <- lmer(formula(paste0("just_educ~","1 + (1|idencuesta)+(1|comuna)")),data = df_study1)
educ1 <- lmer(formula(paste0("just_educ~",h1b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ2 <- lmer(formula(paste0("just_educ~",h2b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ3 <- lmer(formula(paste0("just_educ~",h3b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ4 <- lmer(formula(paste0("just_educ~",h4b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ5 <- lmer(formula(paste0("just_educ~",h5b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ6 <- lmer(formula(paste0("just_educ~",h6b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
educ7 <- lmer(formula(paste0("just_educ~",h7b,"+(1|idencuesta)+(1|comuna)")),data = df_study1)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(educ.null, educ1, educ2, educ3, educ4, educ5, educ6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "Gini comunal",
                                      "Cambio gini 2022-2017",
                                      "Proporción matrícula privada",
                                      "Total matrícula privada",
                                      "Varianza Simce comuna",
                                      "Promedio Simce comuna"
                              )
          #omit.coef = omit
          )
```

## B. Pension distribution justice

```{r distribution-pension}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study1_long_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study1_long_comunas %>%
  select(idencuesta,ola,comuna,just_pension:merit_talent, ahead_family:sexo, ginilag22_17, gini_ypch_2022, gini_ypch, prop_univ, gini_gc, prop_pagado, total_mat_pagado, var_simce_comuna, prom_simce_comuna) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )

# B. Pension distributive justice
pension.null <- lmer(formula(paste0("just_pension~","1 + (1|idencuesta)+(1|comuna)")),data = df_study1)
pension1 <- lmer(formula(paste0("just_pension~",h1,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
pension2 <- lmer(formula(paste0("just_pension~",h2,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
pension3 <- lmer(formula(paste0("just_pension~",h3,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
pension4 <- lmer(formula(paste0("just_pension~",h4,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
pension5 <- lmer(formula(paste0("just_pension~",h5,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
pension6 <- lmer(formula(paste0("just_pension~",h6,"+(1|idencuesta)+(1|comuna)")),data = df_study1)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(pension.null, pension1, pension2, pension3, pension4, pension5, pension6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "ola 2017 <br> <i>(Ref. ola 2016)</i>",
                                      "ola 2018",
                                      "ola 2019",
                                      "ola 2022",
                                      "Merito: esfuerzo",
                                      "Merito: Inteligencia",
                                      "Goahead: Familia",
                                      "Goahead: educacion",
                                      "Goahead: ambicion",
                                      "Goahead: trabajo duro",
                                      "Gini comunal",
                                      "Cambio gini 2022-2017",
                                      "Proporción matrícula privada",
                                      "Total matrícula privada",
                                      "Varianza Simce comuna",
                                      "Promedio Simce comuna",
                                      "Ed. media <br> <i>(Ref. Ed. básica)</i>",
                                      "Ed. técnica",
                                      "Ed. universitaria",
                                      "Quintil 2 <br> <i>(Ref. Quintil 1)</i>",
                                      "Quintil 3",
                                      "Quintil 4",
                                      "Quintil 5",
                                      "Quintil NA",
                                      "Estatus social subjetivo",
                                      "Pos. política: Centro <br> <i>(Ref. izquierda)</i>",
                                      "Pos. política: Derecha",
                                      "Pos. política: No se identifica",
                                      "Edad 30-49 <br> <i>(Ref. 18-29)</i>",
                                      "Edad 50-64",
                                      "Edad 65 o más",
                                      "Mujer <br> <i>(Ref. Hombre)</i>"
                              )
          #omit.coef = omit
          )
```

## C. Health distribution justice

```{r distribution-health}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study1_long_comunas.RData"))

# generate analytical sample
df_study1 <- 
  df_study1_long_comunas %>%
  select(idencuesta,ola,comuna,just_salud:merit_talent, ahead_family:sexo, ginilag22_17, gini_ypch_2022, gini_ypch, prop_univ, gini_gc, prop_pagado, total_mat_pagado, var_simce_comuna, prom_simce_comuna) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )

# C. Health distributive justice
salud.null <- lmer(formula(paste0("just_salud~","1 + (1|idencuesta)+(1|comuna)")),data = df_study1)
salud1 <- lmer(formula(paste0("just_salud~",h1,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
salud2 <- lmer(formula(paste0("just_salud~",h2,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
salud3 <- lmer(formula(paste0("just_salud~",h3,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
salud4 <- lmer(formula(paste0("just_salud~",h4,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
salud5 <- lmer(formula(paste0("just_salud~",h5,"+(1|idencuesta)+(1|comuna)")),data = df_study1)
salud6 <- lmer(formula(paste0("just_salud~",h6,"+(1|idencuesta)+(1|comuna)")),data = df_study1)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(salud.null, salud1, salud2, salud3, salud4, salud5, salud6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "ola 2017 <br> <i>(Ref. ola 2016)</i>",
                                      "ola 2018",
                                      "ola 2019",
                                      "ola 2022",
                                      "Merito: esfuerzo",
                                      "Merito: Inteligencia",
                                      "Goahead: Familia",
                                      "Goahead: educacion",
                                      "Goahead: ambicion",
                                      "Goahead: trabajo duro",
                                      "Gini comunal",
                                      "Cambio gini 2022-2017",
                                      "Proporción matrícula privada",
                                      "Total matrícula privada",
                                      "Varianza Simce comuna",
                                      "Promedio Simce comuna",
                                      "Ed. media <br> <i>(Ref. Ed. básica)</i>",
                                      "Ed. técnica",
                                      "Ed. universitaria",
                                      "Quintil 2 <br> <i>(Ref. Quintil 1)</i>",
                                      "Quintil 3",
                                      "Quintil 4",
                                      "Quintil 5",
                                      "Quintil NA",
                                      "Estatus social subjetivo",
                                      "Pos. política: Centro <br> <i>(Ref. izquierda)</i>",
                                      "Pos. política: Derecha",
                                      "Pos. política: No se identifica",
                                      "Edad 30-49 <br> <i>(Ref. 18-29)</i>",
                                      "Edad 50-64",
                                      "Edad 65 o más",
                                      "Mujer <br> <i>(Ref. Hombre)</i>"
                              )
          #omit.coef = omit
          )
```

## D. Income differences are too large

```{r inequality}
#rm(list=ls())
#if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               summarytools,
               lme4,
               texreg
               )
load(file = here::here("input/data-proc/df_study2_long_comunas.RData"))

# generate analytical sample
df_study2 <- 
  df_study2_long_comunas %>%
  select(idencuesta,ola,comuna,des_perc:merit_talent, ahead_family:sexo, ginilag22_17, gini_ypch_2022, gini_ypch, prop_univ, gini_gc, prop_pagado, total_mat_pagado, var_simce_comuna, prom_simce_comuna) %>% 
  na.omit() %>% 
  mutate(ola=as.factor(ola),
         sexo=as.factor(sexo)
         )

# C. Health distributive justice
des.null <- lmer(formula(paste0("des_perc~","1 + (1|idencuesta)+(1|comuna)")),data = df_study2)
des1 <- lmer(formula(paste0("des_perc~",h1,"+(1|idencuesta)+(1|comuna)")),data = df_study2)
des2 <- lmer(formula(paste0("des_perc~",h2,"+(1|idencuesta)+(1|comuna)")),data = df_study2)
des3 <- lmer(formula(paste0("des_perc~",h3,"+(1|idencuesta)+(1|comuna)")),data = df_study2)
des4 <- lmer(formula(paste0("des_perc~",h4,"+(1|idencuesta)+(1|comuna)")),data = df_study2)
des5 <- lmer(formula(paste0("des_perc~",h5,"+(1|idencuesta)+(1|comuna)")),data = df_study2)
des6 <- lmer(formula(paste0("des_perc~",h6,"+(1|idencuesta)+(1|comuna)")),data = df_study2)

#omit <- "(pos_id)|(edad)|(cuestion_mig)|(nation_t1)|(quintil1QNA)|(Intercept)"
texreg::htmlreg(list(des.null, des1, des2, des3, des4, des5, des6),
                custom.note = "*** p < 0.001; ** p < 0.01; * p < 0.05", 
                custom.coef.names = c("Intercepto",
                                      "ola 2017 <br> <i>(Ref. ola 2016)</i>",
                                      "ola 2018",
                                      "ola 2019",
           #                           "Ola 2021",
                                      "ola 2022",
                                      "Merito: esfuerzo",
                                      "Merito: Inteligencia",
                                      "Goahead: Familia",
                                      "Goahead: educacion",
                                      "Goahead: ambicion",
                                      "Goahead: trabajo duro",
                                      "Gini comunal",
                                      "Cambio gini 2022-2017",
                                      "Proporción matrícula privada",
                                      "Total matrícula privada",
                                      "Varianza Simce comuna",
                                      "Promedio Simce comuna",
                                      "Ed. media <br> <i>(Ref. Ed. básica)</i>",
                                      "Ed. técnica",
                                      "Ed. universitaria",
                                      "Quintil 2 <br> <i>(Ref. Quintil 1)</i>",
                                      "Quintil 3",
                                      "Quintil 4",
                                      "Quintil 5",
                                      "Quintil NA",
                                      "Estatus social subjetivo",
                                      "Pos. política: Centro <br> <i>(Ref. izquierda)</i>",
                                      "Pos. política: Derecha",
                                      "Pos. política: No se identifica",
                                      "Edad 30-49 <br> <i>(Ref. 18-29)</i>",
                                      "Edad 50-64",
                                      "Edad 65 o más",
                                      "Mujer <br> <i>(Ref. Hombre)</i>"
                              )
          #omit.coef = omit
          )
```

### 2023-12-11 production/prod_prep-mergedata.Rmd
update alternativas y readme
* update analisis alternativos
* update link readme

@@

@@

@@

>>>>

df_study2_long_comunas <- df_study2_long_comunas %>% mutate(year= case_when(ola==1~2017,
                                                                            ola==2~2017,
                                                                            ola==3~2017,
                                                                            ola==4~2022,
                                                                            ola==6~2022))

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long_comunas,
                  y = select(df_gini_comunas_long, comunacod=COMUNA_COD,year, starts_with("gini"),prop_univ), 
                  by = c("comunacod", "year"))

@@

df_study2_long_comunas <- df_study2_long_comunas %>% mutate(year= case_when(ola==1~2016,
                                                                            ola==2~2017,
                                                                            ola==3~2018,
                                                                            ola==4~2019,
                                                                            ola==5~2021,
                                                                            ola==6~2022))

@@

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long_comunas,
                  y = select(df_matricula_comunas_long, comunacod=cod_com_rbd_15,year, total_mat_pagado,prop_pagado), 
                  by = c("comunacod", "year"))

@@

df_study2_long_comunas <- 
 dplyr::left_join(x = df_study2_long_comunas,
                  y = select(df_simce_comunas_long, comunacod=cod_com_15,year, prom_simce_comuna,var_simce_comuna), 
                  by = c("comunacod", "year"))
